---
title: 'Session 2: Homework 1'
author: "Study Group 7"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(dplyr)
library(psych)
library(ggrepel)
library(patchwork)
```

# Rents in San Francsisco 2000-2018

```{r}
# download directly off tidytuesdaygithub repo

rent <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-05/rent.csv')

```

What are the variable types? Do they all correspond to what they really are? Which variables have most missing values?

This dataset has 2 different types of data sets - character and double. All the variables correspond to the respective data types. The character data type is used for all text based variables like city, country, etc and the double precision data type is used for numeric and date variables. 'descr' has the most missing value, 197542. 'adress' has the second most missing value, of 196888.


```{r skim_data}
skimr::skim(rent)
spec(rent)
unique(rent$city)
```



Make a plot that shows the top 20 cities in terms of % of classifieds between 2000-2018. You need to calculate the number of listings by city, and then convert that number to a %.

```{r top_cities}
rent %>%
  count(city, sort=TRUE) %>% 
  mutate(prop = n/sum(n)) %>% 
  slice_max(n= 20, order_by= prop) %>% 
  mutate(city = fct_reorder(city,prop)) %>% 
  ggplot(aes(x= prop, y= city))+
  geom_col()+
  theme_minimal(base_size = 18)+
  scale_x_continuous(labels = scales::percent)+labs (x = 'prop', y = 'city', caption = "Source: Kate(2018). Bay Area Craiglist Rental Housing Posts, 2000 - 2018")
```

Make a plot that shows the evolution of median prices in San Francisco for 0, 1, 2, and 3 bedrooms listings. 
```{r}
rent_sf<- rent %>%
          filter(city =="san francisco", beds <= 3) %>%
          group_by(year, beds) %>%
          summarise(med_rn = median(price))
 
ggplot(rent_sf, aes(x=year, y=med_rn, color= factor(beds)))+geom_line()+facet_grid(~beds)
```
```{r}
filtered_beds_0_2018 <- rent %>% group_by(beds,year) %>% filter(city =='san francisco',beds == 0, year ==2018) 
head(filtered_beds_0_2018)
```



Finally, make a plot that shows median rental prices for the top 2 cities in the Bay area. Your final graph should look like this
```{r}
top_twelve_data <- rent %>%
  count(city) %>% #to count number of classifieds in each city
  arrange(desc(n)) %>%
  slice(1:12) 
top_twelve_cities <- as.vector(top_twelve_data$city)
top_twelve_cities

rent %>%
  filter(city %in% top_twelve_cities) %>%
  group_by(year, city) %>%
  summarize(median_price = median(price)) %>%
ggplot(aes(year, median_price, color = factor(city))) +
  geom_line() +
  facet_wrap(vars(city)) +
  labs(
  title = "Rental prices for 1-bedroom flats in the Bay Area",
  x = NULL,
  y = NULL,
  caption = "Source: Pennington. Kate (2018). Bay Area Craigslist Rental Housing Posts 2000-2018") +
  theme(legend.position = "None")
```



The data set shows us that San Francisco has the maximum percentage listings, precisely 4x listings more than San Jose, the second highest listings city. This is valid since most listing are in the densely populated area of San Francisco rather than in the suburbs in the vicinity. Moreover, its vicinity to Silicon Valley makes it a hub for residences. 

In San Francisco, the median prices see a major rise across all the types of bedroom (0-3 bedroom flats) across a 20 year period. The drop in rents from 2002 to 2006 was a result of the housing bubble that caused the demand to decrease and rent prices to plummet down. This bubble burst in 2008, post which we see a massive rise in rents (~90% rise in rent from 2010 to 2015). This similar trend was experienced across all the cities in the Bay Area, with a decline from 2002 to 2006, and a multifold rise in the years after that.

# Analysis of movies- IMDB dataset

```{r,load_movies, warning=FALSE, message=FALSE, eval=FALSE}
movies <- read.csv(here::here("data", "movies.csv"))

```


## Use your data import, inspection, and cleaning skills to answer the following:

-   Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries? 
```{r count_NA, warning=FALSE, message=FALSE, eval=FALSE}
count_NA <- sum(is.na(movies)) #counting all the NAs
count_NA
duplicated_count <- sum(duplicated(movies)) #counting all the duplicated
duplicated_count
```


There are no missing values and duplicated values, if we use skim(movies) to check. 

-   Produce a table with the count of movies by genre, ranked in descending order. We can see that Comedy has the most movies listed, followed by Action. Whereas Thriller only has one movie listed.
```{r table}
movies <- read.csv(here::here("data", "movies.csv"))
movie_table<- movies %>%group_by(genre) %>% count(genre) #grouping by genre, then counting.
arrange(movie_table, desc(n))
```

-   Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many \$ did a movie make at the box office for each \$ of its budget. Ranked genres by this `return_on_budget` in descending order.

We can see that Musical has the highest return on budget at 28.9, followed by Family which had a return on budget at 10.1. Whereas Thriller has the least return on budget at 0.0008.
```{r}
movies %>% 
  group_by(genre) %>% 
  summarize(return_on_budget = (mean(gross)/mean(budget))) %>%  #creating a new column called return_on _budget
  arrange(desc(return_on_budget))
  
  
```

-   Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. 

We can see that Steven Spielberg had a highest sum of gross revenue.
```{r}
movies %>% group_by(director) %>% summarize(sum = sum(gross),mean = mean(gross), median = median(gross), std = sd(gross)) %>%  arrange(desc(sum)) %>% slice_head(n = 15)
```

-   Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.
```{r}
 movies %>% 
  group_by(genre) %>% 
  skim()
  
```


## Use `ggplot` to answer the following

-   Examine the relationship between `gross` and `cast_facebook_likes`. Produce a scatterplot and write one sentence discussing whether the number of facebook likes that the cast has received is likely to be a good predictor of how much money a movie will make at the box office. What variable are you going to map to the Y- and X- axes.

We put "cast_facebook_likes" into the X axis and "gross" to the Y axis. There is a slight positive correlation between 'gross' and 'cast_facebook_like', although there exist many outliers, such as when likes is less than 100. Therefore, the cast is not likely to be a good predictor. However, many data points lies outside the a line of best fit, therefore, we cannot say with good confidence that it is a good predictor.

```{r, gross_on_fblikes}
ggplot(movies, aes(x=cast_facebook_likes, y=gross))+ geom_point() +scale_x_log10() +geom_smooth(method ='lm')
```

-   Examine the relationship between `gross` and `budget`. Produce a scatterplot and write one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.

Budget appears to be a good predictor of grossing, since the graph shows shows a positive correlation. Although when the budget is less than 1 million dollars, it is not a good predictor.

```{r, gross_on_budget}
  ggplot(movies, aes(x=budget, y=gross))+ geom_point() + geom_smooth(method='lm', se=FALSE)+ scale_x_log10() + labs(title='relationship between budget and gross')
```

-   Examine the relationship between `gross` and `rating`. Produce a scatterplot, faceted by `genre` and discuss whether IMDB ratings are likely to be a good predictor of how much money a movie will make at the box office. Is there anything strange in this dataset?

Generally, some genre have much more rating than others,for instance, crime, documentary, and mystery. The others, however, have significantly less reviews. For those with high numbers of reviews, such as Action, Adventure, there exist positive correlation between gross and rating which makes IMDB rating a good predictor for those movies.
```{r, gross_on_rating}
  ggplot(movies, aes(x=rating, y=gross))+ geom_point() +geom_smooth(method='lm')+ scale_x_log10() + facet_wrap(~genre) + theme_minimal() + labs(title = 'Relationship between Rating and Gross for each genre')
```

