---

title: "testtest"
output: html_document
---
# Challenge 1: Excess rentals in TfL bike sharing



```{r, get_tfl_data, cache=TRUE}
url <- "https://data.london.gov.uk/download/number-bicycle-hires/ac29363e-e0cb-47cc-a97a-e216d900a6b0/tfl-daily-cycle-hires.xlsx"

# Download TFL data to temporary file
httr::GET(url, write_disk(bike.temp <- tempfile(fileext = ".xlsx")))

# Use read_excel to read it as dataframe
bike0 <- read_excel(bike.temp,
                   sheet = "Data",
                   range = cell_cols("A:B"))

# change dates to get year, month, and week
bike <- bike0 %>% 
  clean_names() %>% 
  rename (bikes_hired = number_of_bicycle_hires) %>% 
  mutate (year = year(day),
          month = lubridate::month(day, label = TRUE),
          week = isoweek(day))
```


## Monthly Changes in in TfL bike rentals

```{r tfl_absolute_monthly_change, echo=FALSE, out.width="100%"}

bike1<- bike %>% 
  filter (year > 2016, year < 2023) %>% # to obtain years 2017-2022
  group_by(year, month) %>% # group by year, then month
  mutate(month=match(month,month.abb)) %>% # converting month into numbers (to be able to draw)
  summarize(mon_year_avg = mean(bikes_hired)) # find mean of each year of each month

bike2 <- bike %>% 
  filter(year>2015, year <2020) %>% # second line, draw average 2016-2019
  group_by(month) %>%  #
  mutate(month = match(month,month.abb)) %>%  # converting month into numbers (to be able to draw)
  summarize(mean_months = mean(bikes_hired)) # fund mean of each month throughout the years

#mon_year_avg>mean_months #green
#mean_months>mon_year_avg #red
bike3 <- merge(x=bike1,y=bike2,by ='month') #
ggplot(bike3, aes(x=month))+ #
  geom_line(aes(y=mon_year_avg))+ #plot yearly month average
  geom_line(aes(y=mean_months),colour='blue') + #mean of months
  geom_ribbon(data=bike3,aes(ymax=pmin(mean_months,mon_year_avg),ymin=mon_year_avg),fill="green",alpha=0.5) + #green ribbon when average is more then 2016-2019
  geom_ribbon(data=bike3,aes(ymax=pmin(mean_months,mon_year_avg), ymin=mean_months),fill="red",alpha=0.5)+
  #geom_ribbon(data = bike6, aes(ymax = pmax(per_change, 0), ymin = 0), fill = "green", alpha = 0.5)+
  #geom_ribbon(data = bike6, aes(ymax = 0, ymin = pmin(per_change, 0)), fill = "red", alpha = 0.5)
  facet_wrap(vars(year))+ 
  ylim(10000,45000)+ #limit on y axis
  labs(x="",y="Bike rentals",title="Monthly changes in TfL bike rentals",subtitle="Change from monthly average shown in blue and calculated between 2016-2019") +
  theme_minimal()

```

## Weekly Changes in TfL bike rentals
The second one looks at percentage changes from the expected level of
weekly rentals. The two grey shaded rectangles correspond to Q2 (weeks
14-26) and Q4 (weeks 40-52).

```{r tfl_percent_change, echo=FALSE, out.width="100%"}
#process data to get the percentage difference between expected and actual bike rented per week
bike4<- bike %>% filter (year > 2016, year < 2023) %>% group_by(year, week) %>%   summarize(week_year_avg = mean(bikes_hired))

bike5 <- bike %>% filter(year>2015, year <2020) %>% group_by(week)%>% summarize(mean_weeks = mean(bikes_hired), base_line = 0)

bike6 <- merge(x=bike4,y=bike5,by ='week')
bike6<- bike6 %>% filter() %>% mutate(per_change = (week_year_avg - mean_weeks)/mean_weeks, if_exceed = ifelse(per_change >=0, TRUE, FALSE))

#plot expected and actual bike rental per year 
ggplot(bike6, aes(x=week))+geom_line(aes(group = 1, y=per_change),colour='black')+geom_line(aes(y = base_line )) +facet_wrap(vars(year))+xlim(0, 53)+
#setting different colors for above/below 0
geom_ribbon(data = bike6, aes(ymax = pmax(per_change, 0), ymin = 0), fill = "green", alpha = 0.5)+
geom_ribbon(data = bike6, aes(ymax = 0, ymin = pmin(per_change, 0)), fill = "red", alpha = 0.5)+ scale_y_continuous(labels = scales::percent)+
#add geom_rug
geom_rug(mapping = aes(color = if_exceed), sides = "b")+
#add titles
labs(title = "Weekly changes in Tfl bike rental",
              subtitle = "% change from weekly averages calculated between 2016-2019",
              caption = "Data source: TfL, London Data Store")

```

### Should you use the mean or the median to calculate your expected rentals? Why?

We use the mean to calculate the expected rentals. We used the mean to calculate our expected rentals because the TFL may be more inclined to use the data for profit and revenue analysis. Mean, instead of median, may be more accurate to represent the revenue from bike rental.

